---
title: "Live session Unit 05 assignment"
author: "Olga Tanyuk"
date: "February 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Backstory:
Your client is expecting a baby soon.  However, he is not sure what to name the child.  Being out of the loop, he hires you to help him figure out popular names.  He provides for you raw data in order to help you make a decision.

**1. Data Munging**

a. First, import the .txt file into R so you can process it.  Keep in mind this is not a CSV file.  You might have to open the file to see what you’re dealing with.  Assign the resulting data frame to an object, df, that consists of three columns with human-readable column names for each.
```{r}
yob2016_2 <- read.csv("C:/Users/Norman/Desktop/yob2016_2.txt", header=FALSE, sep=";")
df <-yob2016_2
names(df)<-c("first name", "gender", "amount of children")
head(df)
```

b. Display the summary and structure of df
```{r}
summary(df)
str(df)
```

c. Your client tells you that there is a problem with the raw file.  One name was entered twice and misspelled.  The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y’s at the end of the name.  Write an R command to figure out which name it is and display it.
```{r}
grep("yyy$", df$`first name`)
```
Result is 212th name ending with "yyy":
```{r}
df[212,1]
```

d. Upon finding the misspelled name, please remove this particular observation, as the client says it’s redundant.  Save the remaining dataset as an object: y2016
```{r}
y2016 <- df[-212,]
```

**2. Data Merging**
Utilize yob2015.txt for this question.  This file is similar to yob2016, but contains names, gender, and total children given that name for the year 2015.

a. Like 1a, please import the .txt file into R.  Look at the file before you do.  You might have to change some options to import it properly.  Again, please give the dataframe human-readable column names.  Assign the dataframe to y2015. 
```{r}
yob2015_2 <- read.csv("C:/Users/Norman/Desktop/yob2015_2.txt", header=FALSE)
y2015 <-yob2015_2
names(y2015)<-c("first name", "gender", "amount of children")
```

b. Display the last ten rows in the dataframe.  Describe something you find interesting about these 10 rows.
```{r}
tail(y2015, 10)
```
*Interesting thing that we have unusual Male's names starting with "Z" that have the same frequency 5. If it is a mistake we will catch it after merging and will delete those lines.*

c. Merge y2016 and y2015 by your Name column; assign it to final.  The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.
```{r}
final1 <- merge(y2015, y2016, by="first name", all = TRUE)
final2 <- subset(x=final1, !is.na(final1$gender.x))
final <- subset(x=final2, !is.na(final2$gender.y))
```

**3. Data Summary: Utilize your data frame object final for this part.**

a. Create a new column called “Total” in final that adds the amount of children in 2015 and 2016 together.  In those two years combined, how many people were given popular names?
```{r}
amount_of_children <- c(final$`amount of children.x`+final$`amount of children.y`)
final$Total <- paste(amount_of_children)
head(final)
```

b. Sort the data by Total.  What are the top 10 most popular names?
```{r}
final <- final[order(final$Total),]
tail(final)
```
*After sorting we still can not tell the top 10 most popular names, since data was sorted by first digit of the number, not by value. Lets use another way. Lets check the class of Total data.*
```{r}
summary(final)
```
*We see that class is character, instead of numeric. Lets change class of Total column for numeric and Sort data again.*
```{r}
class(final$Total) <- "numeric"
final <- final[order(final$Total),]
tail(final,10)
```
*So the top 10 most popular names are: Isabella, Jacob, William, Mason, Ava, Sophia, Liam, Noah, Olivia, Emma.*

c. The client is expecting a girl!  Omit boys and give the top 10 most popular girl’s names.
```{r}
final_girl1 <- subset(x=final, !final$gender.x=="M")
final_girl <- subset(x=final_girl1, !final_girl1$gender.y=="M")
tail(final_girl,10)
```
*So the top 10 most popular girl’s names are: Emma, Olivia, Sophia, Ava, Isabella, Mia, Charlotte, Abigail, Emily, Harper.*

**4. Upload to Github**

Please see my Home Work at https://github.com/OlhaTanyuk/SMU-MSDS-homework

