---
title: "Live session Unit 04 assignment"
author: "Olga Tanyuk"
date: "February 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### FiveThirtyEight Data

a. Install the fivethirtyeight package using function install.packages("fivethirtyeight")

b. In the listing of Data sets in package ‘fivethirtyeight,’ assign the eighteenth data set to an object ‘df.’
```{r}
library(fivethirtyeight)
data(package = "fivethirtyeight")
df <- college_recent_grads
```
c. Use a more detailed list of the data sets to write out the URL in a comment to the related news story.
```{r}
vignette("fivethirtyeight", package = "fivethirtyeight")
```
And here is the url related to the story "The Economic Guide To Picking A College Major" http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/

d. Using R command(s), give the dimensions and column names of this data frame.
```{r}
# dimensions
dim(df)
# column names
colnames(df)
```
or we can use function str that will provide both answers
```{r}
str(df)
```

### Data Summary

a. Write an R command that gives you the column names of the data frame.  Right after that, write one that counts the number of columns but not rows.  Hint: The number should match one of your numbers in Question 1d for dimensions.
```{r}
colnames(df)
ncol(df)
```

b. Generate a count of each unique major_category in the data frame.  I recommend using libraries to help.  I have demonstrated one briefly in live-session.  To be clear, this should look like a matrix or data frame containing the major_category and the frequency it occurs in the dataset.  Assign it to major_count.
```{r}
library(plyr)
major_count <- count(df, 'major_category')
major_count
```
c. To make things easier to read, put par(las=2) before your plot to make the text perpendicular to the axis.  Make a barplot of major_count.  Make sure to label the title with something informative (check the vignette if you need), label the x and y axis, and make it any color other than grey.  Assign the major_category labels to their respective bar.  Flip the barplot horizontally so that bars extend to the right, not upward.  All of these options can be done in a single pass of barplot(). Note: It’s okay if it’s wider than the preview pane.

```{r}
par(las=2)
major_category <- count(df, 'major_category')
barplot(major_category$freq, names = major_category$major_category, ylab = "Frequency", main = "Popularity of college majors", col="red")
```

d. Write the fivethirtyeight data to a csv file.  Make sure that it does not have row labels.
```{r}
write.csv(college_recent_grads, file = "college_recent_grads.csv", row.names=FALSE)
```

## Codebook
Please visit my GitHub repository SMU MSDS homework: https://github.com/OlhaTanyuk/SMU-MSDS-homework
