---
title: 'MSDS 6306: Doing Data Science – Exploratory Data'
author: "Olga Tanyuk"
date: "March 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background: 
Your organization is responsible for building new VA hospitals in the
mainland of the United States. You are a little overwhelmed by the prospect, not sure
which places need the most help. You decide to begin by exploring healthcare facility
data recorded by the U.S. Government.

### 1. Mental Health Clinics (40%)
a. This data set is a survey of every known healthcare facility that offers mental health services in the United States in 2015. Navigate to https://datafiles.samhsa.gov/study-dataset/national-mental-health-services-survey-2015-n-mhss-2015-ds0001-nid17098 and select the R download. Look through the codebook PDF for an explanation on certain variables. Upon opening the RDA file, the data set should be inserted into your global environment, which you can then reference.

b. Please create code which lists the State abbreviations without their counts, one abbreviation per State value. It does not have to in data frame format. A vector is fine.

```{r}
data <- get(load("C:\\Users\\Norman\\Desktop\\N_MHSS_2015_DS0001_data_r.rda"))
states <- unique(data$LST)
print(states)
```

c. Filter the data.frame from 1A. We are only interested in the Veterans Administration (VA) medical centers in the mainland United States—create a listing of counts of these centers by state, including only mainland locations. Alaska, Hawaii, and U.S.territories should be omitted. DC, while not a state, is in the mainland, so it should remain included. Convert this to data.frame().

```{r}
territories <- grep("AK|HI|AS|GU|PR|VI", data$LST)
data1 <- data[-territories,]
#Lets check that we have now only mainland locations
unique(data1$LST)
States <- data1$LST
States_Freq <- as.data.frame(table(States))
head(States_Freq,5)
Zero <- grep("AK|HI|AS|GU|PR|VI", States_Freq$States)
Final_States_Freq <- States_Freq[-Zero,]
head(Final_States_Freq,5)
```

d. Create a ggplot barchart of this filtered data set. Vary the bar’s colors by what State it has listed. Give it an appropriately professional title that is centered. Make sure you have informative axis labels. The State axis should be readable, not layered over each other. You’re welcome to have a legend or not.

```{r}
library(ggplot2)
mytheme <- theme(plot.title = element_text(face = "bold.italic",size= 14, color = "black"), axis.text.x = element_text(size = 9, face ="bold.italic"), axis.text.y = element_text(face = "bold.italic", size = 7))
ggplot(data = Final_States_Freq, aes(x = Final_States_Freq$States, y = Final_States_Freq$Freq, fill=States), horis=TRUE)+geom_bar(stat = "identity")+labs(title = "Veterans Administration(VA) medical centers in the mainland US", x="States",y="Counts of the VA medical centers")+mytheme+coord_flip()
```

### 2. Cleaning and Bringing in New Features (60%)

a. This graph (1D) might be somewhat misleading, as bigger states may have more hospitals, but could be more sparsely located. Read statesize.csv into your R environment. This contains essentially a vector of square miles for each state. In trying to merge it with your data.frame() from 1C, you find that they don’t match. Use paste() on your LST column in 1C to see what the matter is, and write what you observe in a comment.

b. Correct the problem with the LST column using any method in R that is programmatic and easily understandable. Once you have made these state abbreviations identical to statesize.csv’s Abbrev column, merge the data.frame() from 1C and statesize.csv in order to add size information.

```{r}
statesize <- read.csv("~/R_files/statesize.csv")
#Lets compare variable's class in both data sets
str(Final_States_Freq)
str(statesize)
#Lets rename "Abbrev" for "States" in the "statesize" data set
names(statesize) <- c("StateName","SqMiles","States","Region")
#Lets merge two data sets
MergedData <- merge(Final_States_Freq, statesize, by="States", all=TRUE)
head(MergedData,5)
#Data sets do not match, lets find out why
paste(Final_States_Freq$States)
paste(statesize$States)
#We have extra spaces in the names of States in the Final_States_Freq data set, lets get rid of them
for (i in 1:4) {
    Final_States_Freq$States <- gsub('.{ }$', '',Final_States_Freq$States)
}
paste(Final_States_Freq$States)
#Now we can merge data sets
MergedData <- merge(Final_States_Freq, statesize, by="States", all=TRUE)
#Lets get rid of NAs
Final_MergedData1 <- subset(x=MergedData, !is.na(MergedData$Freq))
Final_MergedData <- subset(x=Final_MergedData1, !is.na(Final_MergedData1$SqMiles))
head(Final_MergedData,10)
```

c. Calculate a new variable in your combined data.frame() which indicates the VA hospitals per thousand square miles.
```{r}
Hospitals_per_ThousandSQmiles <- c((Final_MergedData$Freq*1000)/Final_MergedData$SqMiles)
Final_MergedData$Hospitals_per_ThousandSQmiles <- paste(Hospitals_per_ThousandSQmiles)
head(Final_MergedData,5)
```

d. Create another ggplot which considers the VAs per square thousand miles, rather than just frequency.

• Make sure the State axis is readable, like before. Change the title and axes as appropriate.

• Modify the ggplot syntax to make your bars in descending order (there are StackOverflow topics for this, and I have demonstrated how in Live Coding in prior classes).

• Color-code the bars based on Region (see the merged data.frame)—however, change the color scheme from the default. Any set of colors is fine, so long as it is readable.

• Keep the legend—you should have four regions and therefore four colors.

```{r}
Final_MergedData$Hospitals_per_ThousandSQmiles <- as.numeric(Final_MergedData$Hospitals_per_ThousandSQmiles)
str(Final_MergedData)
mytheme <- theme(plot.title = element_text(face = "bold.italic",size= 14, color = "black"), axis.text.x = element_text(face ="bold.italic"), axis.text.y = element_text(face = "bold.italic", size = 7))
Final_MergedData$States <- factor(Final_MergedData$States, levels = Final_MergedData$States[order(Final_MergedData$Hospitals_per_ThousandSQmiles)])
ggplot(data = Final_MergedData, aes(x = Final_MergedData$States, y = Final_MergedData$Hospitals_per_ThousandSQmiles, fill=Region), horis=TRUE)+geom_bar(stat = "identity")+labs(title = "VA medical centers per thousand sq miles in the mainland US", x="States",y="VA medical centers per thousand sq miles")+mytheme+coord_flip()+scale_fill_hue(l=40) + scale_y_continuous(breaks=seq(0, 65, 5))
```

e. What patterns do you see? By this metric, is there any region that seems relatively high for VA medical centers per thousand square miles? How about low? Given these data, what advice might you give your boss before you start modeling (and why)?

Based on above plot we can say that Northeast Region has the highest value of the VA medical centers per thousand square miles, and West Region has the lowest value. At this point my advice is to build new hospitals in the West Region.